\section*{Acknowledgements}
Many people generously volunteered their time and expertise to advise us on \Wilds.
We are grateful for all of the helpful suggestions and constructive feedback from:
Aditya Khosla,
Andreas Schlueter,
Annie Chen,
Aleksander Madry,
Alexander D’Amour,
Allison Koenecke,
Alyssa Lees,
Ananya Kumar,
Andrew Beck,
Behzad Haghgoo,
Charles Sutton,
Christopher Yeh,
Cody Coleman,
Dan Hendrycks,
Dan Jurafsky,
Daniel Levy,
Daphne Koller,
David Tellez,
Erik Jones,
Evan Liu,
Fisher Yu,
Georgi Marinov,
Hongseok Namkoong,
Irene Chen,
Jacky Kang,
Jacob Schreiber,
Jacob Steinhardt,
Jared Dunnmon,
Jean Feng,
Jeffrey Sorensen,
Jianmo Ni,
John Hewitt,
John Miller,
Kate Saenko,
Kelly Cochran,
Kensen Shi,
Kyle Loh,
Li Jiang,
Lucy Vasserman,
Ludwig Schmidt,
Luke Oakden-Rayner,
Marco Tulio Ribeiro,
Matthew Lungren,
Megha Srivastava,
Nelson Liu,
Nimit Sohoni,
Pranav Rajpurkar,
Robin Jia,
Rohan Taori,
Sarah Bird,
Sharad Goel,
Sherrie Wang,
Shyamal Buch,
Stefano Ermon,
Steve Yadlowsky,
Tatsunori Hashimoto,
Tengyu Ma,
Vincent Hellendoorn,
Yair Carmon,
Zachary Lipton,
and Zhenghao Chen.

The design of the WILDS benchmark was inspired by the Open Graph Benchmark \citep{hu2020open}, and we are grateful to the Open Graph Benchmark team for their advice and help in setting up our benchmark.

This project was funded by an Open Philanthropy Project Award and NSF Award Grant No.~1805310.
Shiori Sagawa was supported by the Herbert Kunzel Stanford Graduate Fellowship.
Henrik Marklund was supported by the Dr.~Tech.~Marcus Wallenberg Foundation for Education in International Industrial Entrepreneurship, CIFAR, and Google.
Sang Michael Xie and Marvin Zhang were supported by NDSEG Graduate Fellowships.
Weihua Hu was supported by the Funai Overseas Scholarship and the Masason Foundation Fellowship.
Sara Beery was supported by an NSF Graduate Research Fellowship and is a PIMCO Fellow in Data Science.
Jure Leskovec is a Chan Zuckerberg Biohub investigator.
Chelsea Finn is a CIFAR Fellow in the Learning in Machines and Brains Program.

We also gratefully acknowledge the support of DARPA under Nos.~N660011924033 (MCS);
ARO under Nos.~W911NF-16-1-0342 (MURI), W911NF-16-1-0171 (DURIP);
NSF under Nos.~OAC-1835598 (CINES), OAC-1934578 (HDR), CCF-1918940 (Expeditions), IIS-2030477 (RAPID); Stanford Data Science Initiative, Wu Tsai Neurosciences Institute, Chan Zuckerberg Biohub, Amazon, JPMorgan Chase, Docomo, Hitachi, JD.com, KDDI, NVIDIA, Dell, Toshiba, and UnitedHealth Group.
