% \clearpage
% \appendix
%\setcounter{page}{1}
% \title{Diverse Weight Averaging for Out-of-Distribution Generalization\\ \normalfont{Supplementary Material}}
% \maketitle

% further details the theoretical and empirical aspects of our work
This supplementary material complements the main paper.
% The Sections follow a similar order than their related Sections in the main paper.
% We also present some additional results.
It is organized as follows:
\begin{enumerate}
    \item \Cref{app:broader_impact} describes the broader impact of our work.
    \item \Cref{app:limit_proof_swad} points out the limitations of existing flatness-based analysis of WA and shows how our analysis solves these limitations.
    \item \Cref{app:proof} details all the proofs of the propositions and lemmas found in our work.
        \begin{itemize}
            \item \Cref{app:wa_loss,app:proof_bvc} derive the bias-variance-covariance-locality decomposition for WA (\Cref{prop:b_var_cov}).
            \item \Cref{app:bias_correlation} establishes the link between bias and correlation shift (\Cref{prop:bias}).
            \item \Cref{app:var_diversity} establishes the link between variance and diversity shift (\Cref{prop:var}).
            \item \Cref{app:proof_wa_ind} compares WA with one of its member (\Cref{lemma:wa_ind}).
        \end{itemize}
    \item \Cref{app:wa_vs_ens_analysis} empirically compares WA to functional ensembling ENS.
    \item \Cref{app:additional_diversity_analysis} presents some additional diversity results on OfficeHome and PACS.
    \item \reb{\Cref{app:valueofm} ablates the importance of the number of training runs}.
    \item \Cref{app:domainbeddetails} describes our experiments on DomainBed and our per-domain results.
    \item \Cref{app:failure_corr_shift} empirically confirms a limitation of WA approaches expected from our theoretical analysis: they do not tackle correlation shift on ColoredMNIST.
    \item \reb{\Cref{app:llr} suggests DiWA's potential when some target data is available for training \cite{kirichenko2022last}}.
\end{enumerate}

\input{sections/appendix/00_app_broader_impact}
\input{sections/appendix/00_app_limit_hessian_swad}
\input{sections/appendix/01_app_proof}
\input{sections/appendix/02_app_dwa_vs_ens}
\input{sections/appendix/04_app_diversity.tex}
\input{sections/appendix/05_app_domainbed.tex}

% DEPRECATED
% \input{sections/appendix/06_app_correlationshift.tex}
% \input{sections/appendix/app_failedidea.tex}
% \input{sections/appendix/app_initialization}
% \input{sections/appendix/app_hessian.tex}
% \input{sections/appendix/app_domaininformation.tex}
