\section{Broader impact statement}
\label{app:broader_impact}

We believe our paper can have several positive impacts.
\textit{First}, our theoretical analysis enables practitioners to know when averaging strategies succeed (under diversity shift, where variance dominates) or break down (under correlation shift, where bias dominates).
This is key to understand when several models can be combined into a production system, or if the focus should be put on the training objective and/or the data.
\textit{Second}, it sets a new state of the art for \ood generalization under diversity shift without relying on a specific objective, architecture or task prior. It could be useful in medicine \cite{Zech2018,degrave2021ai} or to tackle fairness issues related to under-representation \cite{Sagawa2020Distributionally,blodgett2016demographic,barocas2016big}.
\textit{Finally}, DIWA has no additional inference cost; in contrast, functional ensembling needs one forward per member.
Thus, DiWA removes the carbon footprint overhead of ensembling strategies at test-time.

Yet, our paper may also have some negative impacts.
\textit{First}, it requires independent training of several models. It may motivate practitioners to learn even more networks and average them afterwards.
Note that in \Cref{sec:domainbed}, we restricted ourselves to combining only the runs obtained from the standard ERM grid search from DomainBed \cite{gulrajani2021in}.
\textit{Second}, our model is fully deep learning based with the corresponding risks, \eg adversarial attacks and lack of interpretability.
\textit{Finally}, we do not control its possible use to surveillance or weapon systems.
