\section{Proofs}
\label{app:proof}

\subsection{WA loss derivation}
\label{app:wa_loss}
\begin{lemma*}[\ref{lemma:wa_ensembling}]
    Given $\{\theta_m\}_{m=1}^M$ with learning procedures $L_S^M\triangleq\{l_S^{(m)}\}_{m=1}^M$. Denoting $\Delta_{L_S^M}=\max_{m=1}^{M}\left\|\theta_m-\theta_{\text{WA}}\right\|_2$, $\forall (x,y) \in \X \times \Y$:%
    \begin{equation*}
        f_{\text{WA}}(x) = f_{\text{ENS}}(x) + O(\Delta^2_{L_S^M}) \text{~and~} \ell\left(f_{\text{WA}}(x), y\right) = \ell\left(f_{\text{ENS}}(x) , y\right) + O(\Delta_{L_S^M}^2).%
    \end{equation*}%
\end{lemma*}
\begin{proof}
This proof has two components:
\begin{itemize}
    \item to establish the functional approximation, as \cite{izmailov2018}, it performs Taylor expansion of the models' predictions at the first order.
    \item to establish the loss approximation, as \cite{Wortsman2022ModelSA}, it performs Taylor expansion of the loss at the first order.
\end{itemize}
% \footnote{at the first order instead of the second order in \cite{Wortsman2022ModelSA}.}
\paragraph{Functional approximation}
With a Taylor expansion at the first order of the models' predictions \wrt parameters $\theta$:
\begin{align*}
    f_{\theta_m}&=f_{\text{WA}}+\nabla f_{\text{WA}}^\intercal \Delta_m+ O\left(\left\|\Delta_m\right\|_2^2\right)\\
    f_{\text{ENS}}-f_{\text{WA}}&=\frac{1}{M} \sum_{m=1}^{M}\left(\nabla f_{\text{WA}}^\intercal \Delta_m+O\left(\left\|\Delta_m\right\|_2^{2}\right)\right)
\end{align*}
Therefore, because $\sum_{m=1}^{M} \Delta_m=0$,
\begin{equation} \label{eq:diff_fens_fwa}
    f_{\text{ENS}}-f_{\text{WA}} = O\left(\Delta^2\right) \text{~where~}\Delta=\max_{m=1}^{M}\left\|\Delta_m\right\|_2.
\end{equation}
\paragraph{Loss approximation}
With a Taylor expansion at the zeroth order of the loss \wrt its first input and injecting \Cref{eq:diff_fens_fwa}:
\begin{align*}
    \ell\left(f_{\text{ENS}}(x) ; y\right)&=\ell\left(f_{\text{WA}}(x); y\right)+O\left(\left\|f_{\text{ENS}}(x)-f_{\text{WA}}(x)\right\|_2\right) \\
     \ell\left(f_{\text{ENS}}(x) ; y\right) &= \ell\left(f_{\text{WA}}(x); y\right) + O\left(\Delta^2\right).
\end{align*}
\end{proof}
\subsection{Bias-variance-covariance-locality decomposition}

\begin{remark} \label{remark:identical_assumption}
    Our result in \Cref{prop:b_var_cov} is simplified by leveraging the fact that the learning procedures $L_S^M=\{l_S^{(m)}\}_{m=1}^M$ are identically distributed (i.d.). This assumption naturally holds for DiWA which selects weights from different runs with i.i.d. hyperparameters.
    It may be less obvious why it applies to MA \cite{arpit2021ensemble} and SWAD \cite{cha2021wad}.
    It is even false if the weights $\{\theta(l_S^{(m)})\}_{m=1}^M$ are defined as being taken sequentially along a training trajectory, \ie when $0\leq i < j \leq M$ implies that $l_S^{(i)}$ has fewer training steps than $l_S^{(j)}$.
    We propose an alternative indexing strategy to respect the i.d. assumption.
    Given $M$ weights selected by the weight selection procedure, we draw without replacement the $M$ weights, \ie $\theta(l_S^{(i)})$ refers to the $i^{th}$ sampled weights.
    With this procedure, all weights are i.d. as they are uniformly sampled.
    Critically, their WA are unchanged for the two definitions.%
\end{remark}
\label{app:proof_bvc}
\begin{proposition*}[\ref{prop:b_var_cov}]
    Denoting $\bar{f}_S\left(x\right) = \mathbb{E}_{l_S} \left[f\left(x,\theta\left(l_S\right)\right)\right]$, under identically distributed learning procedures $L_S^M\triangleq\{l_S^{(m)}\}_{m=1}^M$, the expected generalization error on domain $T$ of $\theta_{\text{WA}}(L_S^M)\triangleq\frac{1}{M} \sum\nolimits_{m=1}^{M} \theta_m$ over the joint distribution of $L_S^M$ is:%
    \begin{equation*}\tag{\ref{eq:b_var_cov}}
        \begin{aligned}
             \mathbb{E}_{L_S^M}\mathcal{E}_T(\theta_{\text{WA}}(L_S^M)) &= \mathbb{E}_{(x,y)\sim p_T}\Big[\biasb^2(x, y)+\frac{1}{M} \varb(x)+\frac{M-1}{M} \covb(x)\Big] + O(\Deltab^2), \\
             \text{where~}\biasb(x,y)&=y-\bar{f}_S\left(x\right), \\
             \text{and~}\varb(x)&=\mathbb{E}_{l_S}\left[\left(f(x, \theta(l_S)) - \bar{f}_S\left(x\right)\right)^{2}\right], \\
             \text{and~}\covb(x)&= \mathbb{E}_{l_S,l_S'}\left[\left(f(x,\theta(l_S))-\bar{f}_S\left(x\right)\right)\left(f(x,\theta(l_S')))-\bar{f}_S\left(x\right)\right)\right], \\
             \text{and~}\Deltab^2&=\mathbb{E}_{L_S^M}\Delta_{L_S^M}^2 \text{~with~} \Delta_{L_S^M}=\max_{m=1}^{M}\left\|\theta_m-\theta_{\text{WA}}\right\|_2.%
        \end{aligned}%
    \end{equation*}%
    $\covb$ is the prediction covariance between two member models whose weights are averaged.
    The locality term $\Deltab^2$ is the expected squared maximum distance between weights and their average.
\end{proposition*}
\begin{proof}
    This proof has two components:
    \begin{itemize}
        \item it follows the bias-variance-covariance decomposition from \cite{ueda1996generalization,brown2005between} for functional ensembling. It is tailored to WA by assuming that learning procedures are identically distributed.
        \item it injects the obtained equation into \Cref{lemma:wa_ensembling} to obtain the \Cref{prop:b_var_cov} for WA.
    \end{itemize}
    \paragraph{BVC for ensembling with identically distributed learning procedures}
    With $\bar{f}_S\left(x\right) = \mathbb{E}_{l_S} \left[f\left(x,\theta\left(l_S\right)\right)\right]$, we recall the bias-variance decomposition \cite{kohavi1996bias} (\Cref{eq:b_v}):
    \begin{align*}
        \mathbb{E}_{l_S}\mathcal{E}_T(\theta(l_S)) & = \mathbb{E}_{(x,y)\sim p_T}\Big[\biasb(x,y)^{2}+\varb(x)\Big],\\
        \text{where~}\biasb(x,y) & =  \operatorname{Bias}\{f|(x,y)\} = y-\bar{f}_S\left(x\right), \\
        \text{and~}\varb(x) &= \operatorname{Var}\{f|x\} = \mathbb{E}_{l_S}\left[\left(f(x, \theta(l_S)) - \bar{f}_S\left(x\right)\right)^{2}\right]. \\
    \end{align*}
    Using $f_{\text{ENS}}\triangleq f_{\text{ENS}}(\cdot, \{\theta(l_S^{(m)})\}_{m=1}^M) \triangleq \frac{1}{M} \sum_{m=1}^{M} f(\cdot, \theta(l_S^{(m)}))$ in this decomposition yields,
    \begin{equation} \label{eq:b_v_ens}
        \mathbb{E}_{L_S^M}\mathcal{E}_T(\{\theta(l_S^{(m)})\}_{m=1}^M)=\mathbb{E}_{x\sim p_T}\left[\operatorname{Bias}\left\{f_{\text{ENS}} \mid (x,y)\right\}^{2}+\operatorname{Var}\left\{f_{\text{ENS}} \mid x\right\}\right].
    \end{equation}
    As $f_{\text{ENS}}$ depends on $L_S^M$, we extend the bias into:
    \begin{align*}
        \operatorname{Bias}\left\{f_{\text{ENS}} \mid (x,y)\right\}&=y-\mathbb{E}_{L_S^M}\left[\frac{1}{M} \sum_{m=1}^{M} f(x, \theta(l_S^{(m)}))\right]=y-\frac{1}{M} \sum_{m=1}^{M} \mathbb{E}_{l_S^{(m)}}\left[f(x, \theta(l_S^{(m)}))\right]
    \end{align*}
    Under identically distributed $L_S^M\triangleq\{l_S^{(m)}\}_{m=1}^M$,
    $$
        \frac{1}{M} \sum_{m=1}^{M} \mathbb{E}_{l_S^{(m)}}\left[y-f(x, \theta(l_S^{(m)}))\right]=\mathbb{E}_{l_S}\left[y-f(x, \theta(l_S))\right]=\operatorname{Bias}\{f|(x,y)\}.
    $$
    Thus the bias of ENS is the same as for a single member of the WA.
    
    Regarding the variance:
    \begin{align*}
        &\operatorname{Var}\left\{f_{\text{ENS}} \mid x\right\}=\mathbb{E}_{L_S^M}\left[\left(\frac{1}{M} \sum_{m=1}^{M} f(x, \theta(l_S^{(m)}))-\mathbb{E}_{L_S^M}\left[\frac{1}{M} \sum_{m=1}^{M} f(x, \theta(l_S^{(m)}))\right]\right)^{2}\right].\\
        % &=\frac{1}{M^{2}} \sum_{m=1}^{M} \mathbb{E}_{l_S^{(m)}}\left[\left(f(x, \theta(l_S^{(m)}))-\mathbb{E}_{l_S^{(m)}}\left[f(x, \theta(l_S^{(m)}))\right]\right)^{2}\right]+ \\
        % &\quad\quad \frac{1}{M^{2}} \sum_{m} \sum_{m^{\prime} \neq m} \mathbb{E}_{l_S^{(m)}, l_S^{(m^{\prime})}}[\left(f(x, \theta(l_S^{(m)}))-\mathbb{E}_{l_S^{(m)}}\left[f(x, \theta(l_S^{(m)}))\right]\right) \\
        % & \left(f(x, \theta(l_S^{(m^\prime)}))-\mathbb{E}_{l_S^{(m^{\prime})}}\left[f(x, \theta(l_S^{(m^\prime)}))\right]\right)].
    \end{align*}
    Under identically distributed $L_S^M\triangleq\{l_S^{(m)}\}_{m=1}^M$,
    \begin{align*}
        &\operatorname{Var}\left\{f_{\text{ENS}} \mid x\right\} = \frac{1}{M^{2}} \sum_{m=1}^{M} \mathbb{E}_{l_S}\left[\left(f(x, \theta(l_S))-\mathbb{E}_{l_S}\left[f(x, \theta(l_S))\right]\right)^{2}\right]+ \\
        &\quad\quad\quad\quad \frac{1}{M^{2}} \sum_{m} \sum_{m^{\prime} \neq m} \mathbb{E}_{l_S, l_S^\prime}\left[\left(f(x, \theta(l_S))-\mathbb{E}_{l_S}\left[f(x, \theta(l_S))\right]\right)\left(f(x, \theta(l_S^\prime))-\mathbb{E}_{l_S^\prime}\left[f(x, \theta(l_S^\prime))\right]\right)\right] \\
        &= \frac{1}{M} \mathbb{E}_{l_S}\left[\left(f(x, \theta(l_S))-\mathbb{E}_{l_S}\left[f(x, \theta(l_S))\right]\right)^{2}\right] + \\
        &\quad\quad\quad\quad \frac{M-1}{M}\mathbb{E}_{l_S, l_S^\prime}\left[\left(f(x, \theta(l_S))-\mathbb{E}_{l_S}\left[f(x, \theta(l_S))\right]\right)\left(f(x, \theta(l_S^\prime))-\mathbb{E}_{l_S^\prime}\left[f(x, \theta(l_S^\prime))\right]\right)\right] \\
        &=\frac{1}{M} \varb\left(x\right)+\left(1-\frac{1}{M}\right) \covb\left(x\right).
    \end{align*}
    The variance is split into the variance of a single member (divided by $M$) and a covariance term.
    \paragraph{Combination with \Cref{lemma:wa_ensembling}}
    We recall that per \Cref{lemma:wa_ensembling},
    \begin{align*}
        \ell\left(f_{\text{WA}}(x), y\right) = \ell\left(f_{\text{ENS}}(x) , y\right) + O(\Delta_{L_S^M}^2).
    \end{align*}
    Then we have:
    \begin{align*}
        \mathcal{E}_T(\theta_{\text{WA}}(L_S^M))&=\mathbb{E}_{(x,y) \sim p_T}[\ell(f_{\text{WA}}(x),y)]\\
        &=\mathbb{E}_{(x,y) \sim p_T}[\ell(f_{\text{ENS}}(x),y)] + O(\Delta_{L_S^M}^2) = \mathcal{E}_T(\{\theta(l_S^{(m)})\}_{m=1}^M) + O(\Delta_{L_S^M}^2), \\
        \mathbb{E}_{L_S^M} \mathcal{E}_T(\theta_{\text{WA}}(L_S^M))&=\mathbb{E}_{L_S^M}\mathcal{E}_T(\{\theta(l_S^{(m)})\}_{m=1}^M) + O(\mathbb{E}_{L_S^M}[\Delta_{L_S^M}^2]).
    \end{align*}
    We eventually obtain the result:
    \begin{align*}
        % \mathcal{E}_T(\theta_{\text{WA}}(L_S^M)) =
        \mathbb{E}_{L_S^M}\mathcal{E}_T(\theta_{\text{WA}}(L_S^M)) = \mathbb{E}_{(x,y)\sim p_T}\Big[\biasb(x,y)^{2}+\frac{1}{M} \varb(x)+\frac{M-1}{M} \covb(x)\Big] + O(\Deltab^2).
    \end{align*}
\end{proof}%
\input{sections/appendix/proofs/app_bias_corr_inequality}
\input{sections/appendix/proofs/app_var_div}
\input{sections/appendix/proofs/app_wa_vs_member}
